---
title: "Project"
output:
  pdf_document: default
  html_document: default
date: "2023-04-29"
---
#Loading and reading the cleaned csv dataset from the specified directory
```{r}
data<- read.csv("C:/Users/vibar/Downloads/Drug_clean.csv")

#Numerically encoding the Type of Drug column
data$Status <- ifelse(data$Type == "RX", 0, 
               ifelse(data$Type %in% c("OTC"), 1, 2))

head(data)

```

#Checking for Missing and Duplicate Values
```{r}
#Missing Values
if (any(is.na(data))) {
  print("There are missing values in the dataset.")
} else {
  print("There are no missing values in the dataset.")
}

#For Duplicates
if (any(duplicated(data))) {
  print("There are duplicate rows in the dataset.")
} else {
  print("There are no duplicate rows in the dataset.")
}
```
#Descriptive Statistics
For both the continuous variables "Effective and "Satisfaction"
```{r}
# Calculate mean
mean1<- mean(data$Effective)
cat("The mean of the Effectiveness variable is:", mean1, "\n")
mean2<- mean(data$Satisfaction)
cat("The mean of the Satisfaction variable is:", mean2, "\n")

# Calculate median
median1<- median(data$Effective)
cat("The median of the Effectiveness variable is:", median1, "\n")
median2<- median(data$Satisfaction)
cat("The median of the Satisfaction variable is:", median2, "\n")

# Calculate range
range1<- range(data$Effective)
cat("The range of the Effectiveness variable is:", range1, "\n")
range2<- range(data$Satisfaction)
cat("The range of the Satisfaction variable is:", range2, "\n")

# Calculate standard deviation
sd1<- sd(data$Effective)
cat("The sd of the Effectiveness variable is:", sd1, "\n")
sd2<-sd(data$Satisfaction)
cat("The sd of the Satisfaction variable is:", sd2, "\n")
```

#Graphical analysis of the data
Scatter Plot
```{r}
scatter.smooth(x = data$Satisfaction, y = data$Effective, main = "Satisfaction vs Effectiveness")
points(x = data$Satisfaction, y = data$Effective, col = "green", pch = 16)
points(x = data$Satisfaction, y = data$Effective, col = rgb(0, 0, 1, alpha = 0.5), pch = 16)

```
#Graphical analysis of the data
Box-Plot Of the data
```{r}
par(mfrow = c(1,2))
boxplot(data$Effective,main = "Drug Effectiveness", col = "blue")
boxplot(data$Satisfaction,main = "Drug Satisfaction", col = "green")
```
install.packages("moments")
library(moments)

#Graphical analysis of the data
Histogram 
```{r}

#Count for Drug Type
library(ggplot2)
plot <- ggplot(data, aes(x = Type)) +
  geom_bar(fill = "red") +
  labs(title = "Count Plot of Drug Types", x = "Drug Types", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))

# Display the plot
print(plot)

#Histogram for Effective
hist(data$Effective, col = "blue",
     xlab = "Interval", main = "Effectiveness")


#Histogram for Satisfaction
hist(data$Satisfaction, col = "green",
     xlab = "Interval", main = "Satisfaction")

```

#Graphical analysis of the data
Density Plot
```{r}
library(e1071)
par(mfrow=c(1, 3))
plot(density(data$Status), main="Density Plot: Drug Type", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(data$Status), 2)))  
polygon(density(data$Status), col="red")

plot(density(data$Effective), main="Density Plot: Effectiveness", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(data$Effective), 2)))  
polygon(density(data$Effective), col="blue")

plot(density(data$Satisfaction), main="Density Plot: Satisfaction", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(data$Satisfaction), 2)))  
polygon(density(data$Satisfaction), col="green")

```

#Correlation Test between the variables
```{r}
# Between Status and Effectiveness
cor(data$Status,data$Effective)

# Between Status and Satisfaction
cor(data$Status,data$Satisfaction)
```
#Shapiro-Wilk Test
```{r}
# Perform Shapiro-Wilk test for Effective by Status
shapiro.test(data$Effective)

# Perform Shapiro-Wilk test for Satisfaction by Status
shapiro.test(data$Satisfaction)
```

#Kruskal-Wallis Test
```{r}
# Perform Kruskal-Wallis test for Effective by Status
kruskal.test(Effective ~ Status, data = data)

# Perform Kruskal-Wallis test for Satisfaction by Status
kruskal.test(Satisfaction ~ Status, data = data)

```

#Calculating the Skewness and the Kurtosis for the Continous Variables
```{r}
# Calculate skewness
skew <- skewness(data$Satisfaction)
cat("The skewness of the Satisfaction variable is:", skew, "\n")

# Calculate kurtosis
kurt <- kurtosis(data$Satisfaction)
cat("The kurtosis of the Satisfaction variable is:", kurt, "\n")


# Calculate skewness
skew <- skewness(data$Effective)
cat("The skewness of the Effective variable is:", skew, "\n")

# Calculate kurtosis
kurt <- kurtosis(data$Effective)
cat("The kurtosis of the Effective variable is:", kurt, "\n")
```

#Building the Generalized Linear Model
Effectiveness and Satisfaction Combined
```{r}
model1 <- glm(Status ~ Effective + Satisfaction, data = data)
summary(model1)
```
#Proceeding ahead with Mutlinomial Logistic Regression owing to the high AIC by GLM


#Building the Multinomial Logistic Regression Model
Effectiveness and Satisfaction Combined 
```{r}
library(nnet)
model2 <- multinom(Status ~ Effective + Satisfaction, data = data)
summary(model2)

```
#Multinomial Logistic Regression performs better when compared to GLM on Residual Deviance and AIC


#Building the Multinomial Logistic Regression Model
Between Satisfaction and Status
```{r}
library(nnet)
model3 <- multinom(Status ~ Satisfaction, data = data)

summary(model3)

```


##Building the Multinomial Logistic Regression Model
Between Effectiveness and Status
```{r}
library(nnet)
model4 <- multinom(Status ~ Effective, data = data)
summary(model4)
```
